\pkg{Rgbp}: An \proglang{R} Package for Conjugate Gaussian, Poisson, and Binomial Hierarchical Modeling and Frequency Method Checking on Overdispersed Data

### 2015-05-05 jss/article_v2.tex
introduction of paper

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Hyungsuk Tak, Joseph Kelly, Carl Morris} %% comma-separated
\Plaintitle{Rgbp: Hierarchical Modeling and Frequency Method Checking} %% without formatting
\Shorttitle{\pkg{Rgbp}: Hierarchical Modeling and Frequency Method Checking on over dispersed data} %% a short title (if necessary)
%% an abstract and keywords
\Abstract{
\pkg{Rgbp} is an \proglang{R} package that utilizes  approximate Bayesian machinery to fit two-level conjugate hierarchical models on overdispersed Gaussian, Poisson, and Binomial data and evaluates whether the resulting approximate Bayesian interval estimates for random effects meet the nominal confidence levels via frequency coverage evaluation. The data that \pkg{Rgbp} assumes comprise observed sufficient statistic for each group's random effect, such as an average or a proportion of each group, without population-level data.  The approximate Bayesian tool equipped with the adjustment for density maximization \citep{tang2011} produces approximate point and interval estimates for model parameters including second-level variance component, regression coefficients, and each group's random effect. For the Binomial data, the package provides an option to produce  posterior samples of all the model parameters via the acceptance-rejection method. The package provides a quick way to evaluate coverage rates of the resultant Bayesian interval estimates for random effects via a parametric bootstrapping, which we call  ``frequency method checking."  }
%It is found that such Bayesian-frequentist reconciliation allows \pkg{Rgbp} to have attributes desirable from both perspectives, working well in small samples and yielding good coverage probabilities for its interval estimates.
\Keywords{overdispersion, hierarchical model, adjustment for density maximization, repeated sampling coverage evaluation, \proglang{R}}
\Plainkeywords{multilevel model, conjugate hierarchical generalized linear models, frequency method checking, coverage probability, shrinkage, R} 
%% without formatting
%% at least one keyword must be supplied

@@

\section[Introduction]{Introduction}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.
Gaussian, Poisson, or Binomial data from several independent groups sometimes have more variation than the assumed Gaussian, Poisson, or Binomial distributions. To account for the extra-variability, called overdispersion, a two-level hierarchical model regards the first-level mean parameters as random effects that come from an exchangeable or non-exchangeable population-level conjugate prior distribution. The observed data are sufficient statistics for the random effects, such as averages or proportions, obtained from several independent groups. The model can incorporate covariate information of each group via a linear, log-linear, or logistic regression according to the data type, which results in a non-exchangeable conjugate prior distribution.

The package \pkg{Rgbp} fits the two-level models on overdispersed data using the adjustment for density maximization \citep{tang2011}, a generalization of a commonly-used restricted maximum likelihood estimation (REML). 
 and that  The population-level data    of several independent groups are ubiquo
\pkg{Rgbp} is an \proglang{R} package to estimate and validate a two-level model based on  overdispersed  Gaussian, Poisson and Binomial data.  The estimation procedure adopts the adjustment for density maximization \citep{tang2011} for approximate Bayesian point and interval estimates for random effects and the validation involves checking frequency properties of the procedure via repeated sampling (which we call ``frequency method checking"). It is found that even in small samples our procedure yields good frequency properties in comparison to other methods such as Maximum Likelihood Estimation (MLE).

, producing point and interval estimates of second-level variance component, regression coefficients, and random effects.

This package is useful for both frequentists and Bayesians. Bayesians are able to use the package to see results for a non-informative reference prior before and after constructing their own Bayesian hierarchical model and frequentists are provided with a procedure with good repeated sampling properties.

@@

To continue with a maximization-based estimation procedure but to steer clear of aforementioned boundary issues we make use of adjustment for density maximization (ADM) \citep{carl1988, 1997, 2011}. For our purposes we approximate the posterior distribution of a shrinkage factor with a Beta distribution, which allows us to finally obtain estimates of the posterior moments, i.e.,  of $E(B_{j}\vert\boldsymbol{y})$ and $Var(B_{j}\vert\boldsymbol{y})$, where the notation $\boldsymbol{y}$ represents observed data, without the boundary issues that MLE encounters. See \cite{2011}.

>>>>

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Hyungsuk Tak, Joseph Kelly, Carl Morris} %% comma-separated
\Plaintitle{Rgbp: Hierarchical Modeling and Frequency Method Checking} %% without formatting
\Shorttitle{\pkg{Rgbp}: Hierarchical Modeling and Frequency Method Checking on over dispersed data} %% a short title (if necessary)
%% an abstract and keywords
\Abstract{
\pkg{Rgbp} is an \proglang{R} package that utilizes  approximate Bayesian machinery to fit two-level conjugate hierarchical models on overdispersed Gaussian, Poisson, and Binomial data and evaluates whether the resulting approximate Bayesian interval estimates for random effects meet the nominal confidence levels via frequency coverage evaluation. The data that \pkg{Rgbp} assumes comprise observed sufficient statistics for each random effect, such as an average or a proportion, without population-level data.  The approximate Bayesian tool equipped with the adjustment for density maximization produces approximate point and interval estimates for model parameters including second-level variance component, regression coefficients, and random effects. For the Binomial data, the package provides an option to produce  posterior samples of all the model parameters via the acceptance-rejection method. The package provides a quick way to evaluate coverage rates of the resultant Bayesian interval estimates for random effects via a parametric bootstrapping, which we call  ``frequency method checking."  }
%It is found that such Bayesian-frequentist reconciliation allows \pkg{Rgbp} to have attributes desirable from both perspectives, working well in small samples and yielding good coverage probabilities for its interval estimates.
\Keywords{overdispersion, hierarchical model, adjustment for density maximization, repeated sampling coverage evaluation, \proglang{R}}
\Plainkeywords{multilevel model, conjugate hierarchical generalized linear models, frequency method checking, coverage probability, shrinkage, R} 
%% without formatting
%% at least one keyword must be supplied

@@

\section[Introduction]{Introduction}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.
Gaussian, Poisson, or Binomial data from several independent groups sometimes have more variation than the assumed Gaussian, Poisson, or Binomial distributions on the first-level observed data. To account for the extra-variability, called overdispersion, a two-level hierarchical model regards first-level mean parameters as random effects that come from a population-level conjugate prior distribution. The conjugate prior distribution is exchangeable if no covariates are available and non-exchangeable if the model incorporates covariate information of each group via a linear, log-linear, or logistic regression according to the data type.

With an assumption of homogeneity within each group, the observed data are sufficient statistics for the random effects, such as averages or proportions, possibly together with each group's covariate information. This type of  data is common for a biological analysis on litter data, a meta analysis, or small area estimation problems.  For these data, the two-level model that \pkg{Rgbp} assumes can be considered as a conjugate hierarchical generalized linear model \citep{hglm2006, hglm2010} without any fixed effects.

\pkg{Rgbp} takes a Bayesian approach with our special improper hyper-prior distributions on hyper-parameters, the parameters of the conjugate prior distribution. The hyper-prior distributions lead to Stein's harmonic prior known to produce good repeated sampling coverage properties of the Bayesian interval estimates for random effects in a two-level Gaussian hierarchical model \citep{tang2011, morris2012}.  We apply the hyper-prior distributions to Poisson and Binomial hierarchical models. The goal of our hierarchical modeling is to compare random effects among groups by their Bayesian point and interval estimates that have good frequency properties.

When it comes to fitting the model, \pkg{Rgbp} adopts the adjustment for density maximization \citep{carl1988, 1997, tang2011} (ADM), a generalization of a commonly-used restricted maximum likelihood method (REML). The ADM uses Beta distributions to approximate the posterior distributions of shrinkage factors, functions of a second-level variance component, via maximization like a Delta method that uses a Normal distribution to obtain the approximate distribution of a function of a parameter with its MLE and observed information plugged-in.  The ADM is more appropriate than the Delta method (or REML) because the support of each shrinkage factor is between 0 and 1, not ($-\infty,~\infty$). Also, the ADM is free of boundary effects that REML suffers from (e.g. a zero estimate for the second-level variance component). Using the approximate estimates for the posterior mean and variance of the shrinkage factors, \pkg{Rgbp} obtains  approximate posterior mean and variance of the random effect. Finally, \pkg{Rgbp} approximates the posterior distributions of random effects by a Pearson family distribution whose two parameters are matched to the estimated posterior mean and variance of random effects.

For the Gaussian hierarchical model, \pkg{Rgbp} approximates the posterior distribution of the random effects by a skewed Normal distribution whose three parameters are matched to the first three posterior moment estimates for random effects. For the Binomial  model, \pkg{Rgbp} provides an option to draw independent posterior samples of all the model parameters including second-level variance component, logistic regression coefficients, and random effects,  via an acceptance-rejection method instead of the approximate tool.

In addition to fitting hierarchical models, \pkg{Rgbp} provides a quick way to evaluate the repeated sampling coverage rates of the resulting Bayesian interval estimates for random effects. It is a unique procedure that distinguishes \pkg{Rgbp}  from any other \proglang{R} packages for similar multilevel models, such as \pkg{hglm} for conjugate hierarchical generalized models, \pkg{arm} for Bayesian hierarchical models, and \pkg{hbsae} for hierarchical Bayesian small area estimation. The evaluation procedure which we call ``frequency method checking'' adopts a parametric bootstrapping that generates mock data sets given hyper-parameter values and estimates the coverage rates.

%This package is useful for both frequentist and Bayesian. Bayesians are able to use the package to see results for a non-informative reference prior before and after constructing their own Bayesian hierarchical model and frequentists are provided with a procedure with good repeated sampling properties.

@@

To continue with a maximization-based estimation procedure but to steer clear of aforementioned boundary issues we make use of adjustment for density maximization (ADM) \citep{carl1988, 1997, tang2011}. For our purposes we approximate the posterior distribution of a shrinkage factor with a Beta distribution, which allows us to finally obtain estimates of the posterior moments, i.e.,  of $E(B_{j}\vert\boldsymbol{y})$ and $Var(B_{j}\vert\boldsymbol{y})$, where the notation $\boldsymbol{y}$ represents observed data, without the boundary issues that MLE encounters. See \cite{2011}.
